---
title: "Microbiota adaptation review - 2. Analysis"
author: "Garazi Martin Bideguren"
date: "01-11-2023"
output: 
  pdf_document:
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Review information

Analysis of the papers that met the matching criteria:
a) covers gut microbiota, b) covers environmental adaptations, c) focuses on wildlife/wild animals and d) it is an empirical study./


## Data preparation

### Load required libraries

```{r libraries, echo=TRUE, message=FALSE, results=FALSE }
library(dplyr)
library(tidyverse)
library(devtools)
library(ggplot2)
library(vegan)
```

## Declare directories and files

```{r directories}
workingdir="/Users/dlz554/Documents/Microbiome_adaption_review/Search01.06.23"
R_view="Database-R_view.csv"
weights="Scores_weight.csv"
maxscores="maxscores.csv"
metrics <- colnames(weights)[-1]
```

## Load data

```{r loaddata}
setwd(workingdir)
data <- read.table(R_view,header=T, sep=",")
weights <- read.table(weights,header=T, sep=",")
maxscores <- read.table(maxscores,header=T, sep=",")
```


## Proportional data
Transform metrics to the 0-1 scale.
```{r total_score}
data_proportional <- data %>%
  mutate(across(all_of(metrics), ~ . / maxscores[[(cur_column())]]))
```

## Total score
Compute total scores given a vector of weights.
```{r total_score_1}
weight_vector=weights[2,-1]
data_proportional <- data_proportional %>%
  mutate(across(all_of(metrics), ~ . * weight_vector[[(cur_column())]])) %>%
  mutate(total = rowSums(across(all_of(metrics))) / sum(weight_vector))
```

## Data visualization
### Summary statistics
#### Total scores distribution
Distribution of total scores of the 109 papers analysed
```{r histogram_score, echo=TRUE}
data_proportional %>%
  ggplot(aes(x=Total)) +
  geom_histogram(binwidth = 0.1,) +
  theme_classic()
```

#### Papers analysed per year
The 109 papers analysed distributed per year
```{r histogram_year, echo=TRUE}
data_proportional %>%
  ggplot(aes(x=Year)) +
  geom_histogram(binwidth = 1) +
  theme_classic()
```

#### Total scores distribution in each continent
```{r boxplot_continent, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Continent, sep = ", ") %>%
  ggplot(aes(x = Continent, y = Total, fill = Continent)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape=8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA"))+
  theme_classic() +
  labs(title = "Total Score Distributions by Continent",
       y = "Total Score",
       x = "Continent") +
  coord_flip()
```

#### Total scores distribution in each animal class
```{r boxplot_class, echo=TRUE, warning=FALSE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional$Class <- str_replace(data_proportional$Class, "Dipnoi", "Fish")
data_proportional$Class <- str_replace(data_proportional$Class, "Actinopterygii", "Fish")
data_proportional %>%
  separate_rows(Class, sep = ", ") %>%
  ggplot(aes(x = Class, y = Total, fill = Class)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape=8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677")) +
  scale_colour_manual(values = c("#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677"))+
  theme_classic() +
  labs(title = "Total Score Distributions by Animal Class",
       y = "Total Score",
       x = "Class") +
  coord_flip()
```


#### Total scores distribution per year
```{r boxplot_year, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Total Score Distributions per Year",
       y = "Total Score",
       x = "Year") +
  coord_flip()
```

#### Total scores trend
Total scores trend fitted by a linear model 
```{r scatterplot_year, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```
```{r linear_model_scatterplot_year, echo=TRUE}
linear_model <- lm(Total ~ Year, data = data_proportional)
summary(linear_model)
```


### Total scores of each of the 109 papers divided by the 3 domains
```{r stack_barplot, echo=TRUE}
data_long <- data_proportional %>%
  select(Paper_ID, Year, Reproducibility_stackbarplot, Experimental_stackbarplot, Methodology_stackbarplot) %>%
  pivot_longer(cols = c(Reproducibility_stackbarplot, Experimental_stackbarplot, Methodology_stackbarplot), 
               names_to = "Category", values_to = "Score")
data_long <- data_long %>%
  arrange(Year) %>%
  mutate(Paper_ID = factor(Paper_ID, levels = unique(Paper_ID)))  # Reorder and group by year


ggplot(data_long, aes(x = Paper_ID, y = Score, fill = Category)) +
  geom_bar(stat = "identity") +
  labs(title = "Total scores",
       x = "Paper ID",
       y = "Score") +
  theme_minimal() +
  facet_grid(~ Year, scales = "free", space = "free")+
  theme(axis.text.x = element_blank(),axis.ticks.x=element_blank())+
  scale_fill_discrete(name = "Domain", labels = c("Experimental design", "Methodological resolution", "Reproducibility"))
```

## Criterion scores visualization

#### Experimental approach scores distribution per year
```{r experiment_year, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total_experimental_design_approach, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Experimental approach scores distribution per year",
       y = "Experimental approach scores",
       x = "Year") +
  coord_flip()
```

#### Experiment approach scores trend
```{r experiment_year_trend, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total_experimental_design_approach)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```



#### Fitness measurement scores distribution per year
```{r fitness_year, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total_fitness_measurement, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Fitness measurement scores distribution per year",
       y = "Fitness measurement scores",
       x = "Year") +
  coord_flip()
```

#### Fitness measurement scores trend
```{r fitness_year_trend, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total_fitness_measurement)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```

### Microbiota-host analysis
```{r analysis, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total_microbita_host_analysis, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Microbiota-host analysis",
       y = "Total_microbita_host_analysis",
       x = "Year") +
  coord_flip()
```

```{r analysis_trend, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total_microbita_host_analysis)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```

#### Molecular technique scores distribution per year
```{r technique_year, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total_methodological_resolution, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Molecular technique scores distribution per year",
       y = "Molecular technique scores",
       x = "Year") +
  coord_flip()
```

#### Molecular technique scores trend
```{r techniquet_year_trend, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total_methodological_resolution)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```


### Conclusiveness scores distribution

```{r adaptation_means, echo=TRUE}
adaptation_hyp<-data_proportional %>%
  select(Paper_ID,Year, Adaptation_means,Adaptation_discussion)

ggplot(adaptation_hyp, aes(x = Adaptation_means, fill = Adaptation_discussion)) +
  geom_bar() +
  labs(title = "Adaptation as hypothesis",
       x = "How has been tested",
       y = "Nº of papers") +
  theme_minimal() +
  scale_x_discrete(name ="Aim of the study", labels=c("Descriptive comparison" = "Doesn't test adaptation", "Comparison under host biological process" = "To test adaptation")) +
  theme(axis.text.x = element_text(angle=90))
```

### Conclusiveness scores against total scores obtained

```{r adaptation_means_scores, echo=TRUE}
data_proportional %>%
  ggplot(aes(x=Adaptation_discussion_score,y=Total)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```



## 3 Domain scores visualization
### Experimental design
```{r experimental_design, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total_experimental_design, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Experimental design",
       y = "Total_experimental_design",
       x = "Year") +
  coord_flip()
```

```{r experimental_design_trend, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total_experimental_design)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```

```{r linear_model_scatterplot_experimental_design, echo=TRUE}
linear_model <- lm(Total_experimental_design ~ Year, data = data_proportional)
summary(linear_model)
```




### Methodological resolution
```{r methodology, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total_methodology, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Methodological resolution",
       y = "Total_methodology",
       x = "Year") +
  coord_flip()
```

```{r methodology_trend, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total_methodology)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```

```{r linear_model_scatterplot_methodology, echo=TRUE}
linear_model <- lm(Total_methodology ~ Year, data = data_proportional)
summary(linear_model)
```


#### Data availability scores distribution per year
```{r reproducibility_year, echo=TRUE}
data_proportional$Year<-as.character(data_proportional$Year)
data_proportional %>%
  separate_rows(Year, sep = ", ") %>%
  ggplot(aes(x = Year, y = Total_reproducibility, fill = Year)) +
  geom_violin() +
  geom_point(stat = "summary", fun = "median", shape = 8, color = "black", size = 3) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.5,
    size = 1.5) +
  scale_fill_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  scale_colour_manual(values = c("#E69F00", "#AA3377", "#228833", "#66CCEE", "#CCBB44", "#EE6677", "#4477AA", "#125A56")) +
  theme_classic() +
  labs(title = "Reproducibility scores distribution per year",
       y = "Reproducibility scores",
       x = "Year") +
  coord_flip()
```

#### Data availability scores trend
```{r reproducibility_year_trend, echo=TRUE}
data_proportional$Year<-as.integer(data_proportional$Year)
data_proportional %>%
  ggplot(aes(x=Year,y=Total_reproducibility)) +
  geom_point() +
  geom_smooth(method=lm) +
  theme_classic()
```

```{r linear_model_scatterplot_reproducibility, echo=TRUE}
linear_model <- lm(Total_reproducibility ~ Year, data = data_proportional)
summary(linear_model)
```

